Biologically-Constrained Region Merging for Connectome Reconstruction

### 2018-03-14 papers/eccv2018/2_method.tex
Updates.

\begin{figure}[t]
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_error.png}
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_success.png}
	\end{minipage}
	\caption{Two example candidates for merging. The two segments in the left example do not belong to the same neuron as evidenced by the sharp turning radius.}
	\label{fig:turn-radii}
\end{figure}
\subsection{Learning-based Edge Weight}
\label{sec:edge-weights}
We assign an edge weight $w_e$ to each edge corresponding to the probability that two nodes belong to the same neuronal process.
We train a 3D CNN model to learn these geometric constraints for valid connection between two segments from labeled volume.

\noindent\textbf{Edge Probability}
To predict the probabilities that two segments belong to the same neuron, we train a feed-forward convolutional network with three \textit{VGG-style} convolution blocks~\cite{chatfield2014return} and two fully connected layers before the final sigmoid activation. 
Each \textit{VGG-style} convolution block consists of two double convolutions with filter size $3\times3\times3$ and LeakyReLU(0.0005), followed by a max pooling with filter size $2\times2\times1$ in the first two blocks and $2\times2\times2$ in the third.

For the input of the network, we extract a cubic region of interest (ROI) around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. 
The CNN receives three input channels for every voxel in the ROI around segments $l_1$ and $l_2$. 
The input in all of the channels is in the range $\{-0.5, 0.5\}$. 
The first channel is $0.5$ only if the corresponding voxel has label $l_1$. 
The second channel is $0.5$ only if the corresponding voxel has label $l_2$. 
The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.
We do not use the raw EM image information to avoid the need to retrain the network on datasets that have been stained differently or imaged at different resolution. 
This reduces the need for generating costly manually-labeled ground truth.

\noindent\textbf{Edge Weight}

Given the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e$ is defined as
%p_e &=\mbox{CNN}(x_1, x_2) \\
\begin{align}
w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}}
\end{align}
, where $\beta$ is a tunable parameter that encourages over- or under-segmentation. 
This particular weighting scheme allows us to optimize for a Bayesian network defined on the graph~\cite{keuper2015efficient,andres2011probabilistic}.

@@

After graph construction, we segment the graph into a globally consistent manner while enforcing topological constraints on the output.

%\subsubsection{Lifted Multicut}
\noindent\textbf{Lifted Multicut}
After constructing the graph we seek to partition it into labels where every label corresponds to a neuronal process. 
We formulate this graph partitioning problem as a multicut problem.
There are two primary benefits to using a multicut formulation. 
First, the number of segments in the final graph is not predetermined but depends on the input. 
Second, this minimization produces globally consistent solutions (i.e. a boundary remains only if the two corresponding nodes belong to unique segments)~\cite{keuper2015efficient}.

We apply the algorithms of Keuper et al.~\cite{keuper2015efficient} to produce a feasible solution to the multicut problem using greedy additive edge contraction.
Following their example, we employ the generalized ``lifted'' multicut formulation.
Traditional multicut solutions only consider the probabilities that two adjacent nodes belong to the same segment. 
In the ``lifted'' extension to the problem, we can penalize non-adjacent nodes that belong to different segments. 
These penalties between non-adjacent nodes are called ``lifted'' edges. 
Ideally these ``lifted'' weights represent the probability that two nodes belong to the same neuron.
However, determining such probabilities is computationally expensive.
We approximate these probabilities by finding the maximal probable path between any two nodes using Dijkstra's algorithm~\cite{keuper2015efficient}.
This is an underestimate of the probability that two nodes belong to the same neuron since it does not consider all possible paths.
Since our graphs are sufficiently small, we can generate ``lifted'' edges between all pairs of nodes.

%\subsubsection{Topological Constraints}
\noindent\textbf{Topological Constraints}
By reformulating the segmentation problem as a graph partitioning one, we can enforce some global constraints on our result based on the underlying biology.
Traditional hierarchical clustering algorithms do not rely on such constraints but consider local decisions independently.
We enforce a global constraint that neurons are tree-structured and should not contain cycles. 
The multicut problems returns a series of ``collapsed'' edges between nodes that belong to the same neuron.
We iterate over these edges in order of the probability of merge generated by our CNN. 
We ``collapse'' an edge only if it does not create a cycle in the graph.

>>>>

\begin{figure}[t]
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_error.png}
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_success.png}
	\end{minipage}
	\caption{Two example candidates for merging. The two segments in the left example do not belong to the same neuron as evidenced by the sharp turning radius.}
	\label{fig:turn-radii}
\end{figure}
\subsection{Learning-based Edge Weight}
\label{sec:edge-weights}
We assign an edge weight $w_e$ to each edge corresponding to the probability that two nodes belong to the same neuronal process.
We train a 3D CNN model to learn these geometric constraints for valid connection between two segments from labeled volume.
\\~\\
\noindent\textbf{Edge Probability}
To predict the probabilities that two segments belong to the same neuron, we train a feed-forward convolutional network with three \textit{VGG-style} convolution blocks~\cite{chatfield2014return} and two fully connected layers before the final sigmoid activation.

For the input of the network, we extract a cubic region of interest (ROI) around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. 
The CNN receives three input channels for every voxel in the ROI around segments $l_1$ and $l_2$. 
The input in all of the channels is in the range $\{-0.5, 0.5\}$. 
The first channel is $0.5$ only if the corresponding voxel has label $l_1$. 
The second channel is $0.5$ only if the corresponding voxel has label $l_2$. 
The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.
We do not use the raw EM image information to avoid the need to retrain the network on datasets that have been stained differently or imaged at different resolution. 
This reduces the need for generating costly manually-labeled ground truth. 
\\~\\
\noindent\textbf{Edge Weight} Given the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e$ is defined as
%p_e &=\mbox{CNN}(x_1, x_2) \\
\begin{align}
w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}},
\end{align}
where $\beta$ is a tunable parameter that encourages over- or under-segmentation. 
This particular weighting scheme allows us to optimize for a Bayesian network defined on the graph~\cite{keuper2015efficient,andres2011probabilistic}.

@@

After graph construction, we segment the graph into a globally consistent manner while enforcing topological constraints on the output.
\\~\\
%\subsubsection{Lifted Multicut}
\noindent\textbf{Lifted Multicut}
After constructing the graph we seek to partition it into labels where every label corresponds to a neuronal process. 
We formulate this graph partitioning problem as a multicut problem.
There are two primary benefits to using a multicut formulation. 
First, the number of segments in the final graph is not predetermined but depends on the input. 
Second, this minimization produces globally consistent solutions (i.e. a boundary remains only if the two corresponding nodes belong to unique segments)~\cite{keuper2015efficient}.

We apply the algorithms of Keuper et al.~\cite{keuper2015efficient} to produce a feasible solution to the multicut problem using greedy additive edge contraction.
Following their example, we employ the generalized ``lifted'' multicut formulation.
Traditional multicut solutions only consider the probabilities that two adjacent nodes belong to the same segment. 
In the ``lifted'' extension to the problem, we can penalize non-adjacent nodes that belong to different segments. 
These penalties between non-adjacent nodes are called ``lifted'' edges. 
Ideally these ``lifted'' weights represent the probability that two nodes belong to the same neuron.
However, determining such probabilities is computationally expensive.
We approximate these probabilities by finding the maximal probable path between any two nodes using Dijkstra's algorithm~\cite{keuper2015efficient}.
This is an underestimate of the probability that two nodes belong to the same neuron since it does not consider all possible paths.
Since our graphs are sufficiently small, we can generate ``lifted'' edges between all pairs of nodes. 
\\~\\
%\subsubsection{Topological Constraints}
\noindent\textbf{Topological Constraints}
By reformulating the segmentation problem as a graph partitioning one, we can enforce some global constraints on our result based on the underlying biology.
Traditional hierarchical clustering algorithms do not rely on such constraints but consider local decisions independently.
We enforce a global constraint that neurons are tree-structured and should not contain cycles. 
The multicut problems returns a series of ``collapsed'' edges between nodes that belong to the same neuron.
We iterate over these edges in order of the probability of merge generated by our CNN. 
We ``collapse'' an edge only if it does not create a cycle in the graph.
